---
title: "Joins and strings and dates"
format: html
editor: visual
---

```{r}
library(tidyverse)
library(palmerpenguins)
library(lubridate)
library(kableExtra)
```

## Joins

```{r}
animals <- tibble::tribble(
             ~location,   ~species,  ~maturity,
              "lagoon",   "bobcat",    "adult",
               "bluff",   "coyote", "juvenile",
               "creek",      "fox",    "adult",
                "oaks", "squirrel", "juvenile",
               "bluff",   "bobcat",    "adult"
             )

sites <- tibble::tribble(
           ~location,      ~full_site_name, ~jurisdiction,
             "beach",       "Goleta Beach",     "SB City",
            "lagoon",        "UCSB Lagoon",        "UCSB",
             "bluff",       "Ellwood Mesa",     "SB City",
              "oaks", "Fremont Campground",        "USFS"
           )

```

Full join

```{r}
full_join_example <- full_join(animals, sites)
```

Left join

```{r}
left_join_example <- left_join(animals, sites, by = "location")
```

Inner join

```{r}
inner_join_example <- inner_join(animals, sites, by = "location")
```

```{r}
sites_2 <- sites |> 
  rename(loc = location)
```

Example of joining when key variables have non-matching names.

```{r}
nonmatching <- full_join(animals, sites_2, by = c("location" = "loc"))
```

Semi join

```{r}
semi_join_example <- semi_join(animals, sites, by = "location")
```

Anti join

```{r}
anti_join_example <- anti_join(animals, sites, by = "location")
```

# **Part 1: Data wrangling refresher**

**Refresher 1:** Starting with the `penguins` dataset in the `palmerpenguins` package, write a single piped sequence in which you:

1.  Only include penguins at Biscoe and Dream islands

2.  Remove the `year` and `sex` variables

3.  Add a new column called `body_mass_kg`, with penguin mass converted from grams to kilograms

4.  Rename the `island` variable to `location`

```{r}
penguins_subset <- penguins |> 
  filter(island %in% c("Biscoe", "Dream")) |> 
  select(-year, -sex) |> 
  mutate(body_mass_kg = body_mass_g / 1000) |> 
  rename(location = island)
```

Dates with `lubridate`
```{r}
my_date <- "03-15-1998"
#mdy, ymd, dmy
lubridate::mdy(my_date)

my_date <- "11/7/2003"
lubridate::dmy(my_date)

my_date <- "08-Jun-1974"
lubridate::dmy(my_date)

my_date <-"19610518"
my_real_date <- lubridate::ymd(my_date)
```

Breaking lubridate
```{r}
lubridate::ymd("1942-08-30")

lubridate::mdy("09/22/84")
```

```{r}
time <- "2020-08-12 11:18"
time <- ymd_hm(time)

time
```

```{r}
with_tz(time, "America/Los_Angeles")
with_tz(time, "Asia/Manila")
with_tz(time, "Pacific/Honolulu")

my_time <- lubridate::ymd_hms(time)
year(my_time)
day(my_time)
month(my_time)
hour(my_time)
minute(my_time)
second(my_time)
```

```{r}
teddy_bday <- lubridate::ymd("20170615")

wday(teddy_bday, label = TRUE)

ruth_bday <- lubridate::ymd("19950701")

wday(ruth_bday, label = TRUE)

week(teddy_bday)
```

